---
title: "hw2_tpd2121_problem 3"
output: html_document
date: "2024-10-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r cleaning}
bakers = 
  read_csv(file = "data/bakers.csv") %>% 
  janitor::clean_names() %>% 
  separate_wider_delim(baker_name, " ", names = c("first_name", "last_name")) %>% 
  arrange(series) %>% 
  select(series, first_name, last_name) %>% 
  unite (id, c(series, first_name), sep = ".")

bakes = 
  read_csv(file = "./data/bakes.csv") %>% 
  janitor::clean_names() %>% 
  rename(first_name = baker) %>% 
  unite(id, c(series, first_name), sep = ".")

results = 
  read_csv(file = "./data/results.csv",
           skip = 2) %>% 
  janitor::clean_names() %>% 
  rename(first_name = baker) %>% 
  unite (id, c(series, first_name), sep = ".")

#checking data
anti_join(bakers, bakes)
anti_join(bakers, results)
anti_join(bakes,results)

#merging
gbb_df1 = 
  full_join(bakers, bakes, by = "id")

gbb_df2 = 
  full_join(bakers, results, by = "id") %>% 
  drop_na(result)

gbb_df3 = 
  full_join(gbb_df1, gbb_df2, by = c('id' = 'id', 'last_name' = 'last_name', 'episode' = 'episode'))

#exporting
write.csv(gbb_df3, ".data/gbb_df3.csv")
```

The data cleaning process for all three datasets are quite similar. After importing the dataset into R, I clean the names so all the syntax are uniform then I change the column containing the bakers' name. For the `bakers` dataframe, I separate the bakers' name into first and last names. For `bakes` and `results`, I just changed the column name from `bakers` to `first_name` so it's uniform with the column on the `bakers` dataframe. `bakers` dataframe has additional process of moving the seris to the first column, selecting only the first 3 columns so it contains only `series`, `first_name`, and last_name columns.

For all dataframe, I merged the `series` and `first_name` columns, forming a new variable called `id`. This is done to make it easier in merging, since this variable exists in all dataframe. While checking the data using the `anti_join` function, I realized that season 2 contestant, Joanne Wheatley, is recorded with different `first_name` in each dataframe. Further, the `bakers` and `results` dataframe contains data until season 10, but `bakes` data frame only has data until season 8. 

I used 2 steps in merging the data since it's easier for my understanding (but maybe not the best work practice). The first step is by merging `bakers` and `bakes`, forming `gbb_df1`, then I merged `bakers` and `results` forming `gbb_df2`. For the latter, I also use `drop_na` as eliminated participants will have no relevance in the rest of the series. Finally, I merged `gbb_df1` and `gbb_df2` to form the final data, `gbb_df3`. 

```{r table}
gbb_reader = 
  filter(gbb_df3, result %in% c("WINNER", "STAR BAKER")) %>% 
  select(-last_name) %>% 
  separate_wider_delim(id, ".", names = c("series", "first_name")) %>% 
  mutate(
    series = case_match(
      series, 
      "1" ~ "series_01",
      "2" ~ "series_02",
      "3" ~ "series_03",
      "4" ~ "series_04",
      "5" ~ "series_05",
      "6" ~ "series_06",
      "7" ~ "series_07",
      "8" ~ "series_08",
      "9" ~ "series_09",
      "10" ~ "series_10"
      )
  ) %>% 
  select(series, episode, everything()) %>% 
  arrange(series, episode) %>% 
  filter(series %in% c("series_05", "series_06", "series_07", "series_08", "series_09", "series_10")) %>% 
  knitr::kable()

knitr::kable(gbb_reader)
```

Overall, the winners for series are quite predictable as they tend to be the best two in terms of technical in their seasons, and won star baker multiple times. 

```{r viewers}
viewers = 
  read_csv(file = "./data/viewers.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    series_1:series_10,
    names_to = "series",
    values_to = "viewership"
    ) %>% 
  select(series, everything()) %>% 
  mutate(
    series = case_match(
      series, 
      "series_1" ~ "series_01",
      "series_2" ~ "series_02",
      "series_3" ~ "series_03",
      "series_4" ~ "series_04",
      "series_5" ~ "series_05",
      "series_6" ~ "series_06",
      "series_7" ~ "series_07",
      "series_8" ~ "series_08",
      "series_9" ~ "series_09",
      "series_10" ~ "series_10"
      )
  ) %>% 
  arrange(series, episode)

view_1 = 
  filter(viewers, series == "series_01")

view_5 = 
  filter(viewers, series == "series_05")
```

The average viewership for season 1 is `r mean(pull(view_1, viewership), na.rm=TRUE)` for the first 6 episodes with available data, while season 5 is `r mean(pull(view_5, viewership) , na.rm=TRUE)` for all 10 episodes.